---
title: "Instacart"
author: "Bharat"
date: "26 July 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#setwd("E://instacart")
library(data.table)
library(dplyr)
library(markdown)
library(knitr)
library(ggplot2)
library(stringr)
library(DT)
```

#Read in the data
```{r}

orders <-  fread("E:/instacart/orders.csv")
products <-  fread("E:/instacart/products.csv")
order_products <- fread('E:/instacart/order_products__train.csv')
order_products_prior <- fread('E:/instacart/order_products__prior.csv')
aisles <- fread('E:/instacart/aisles.csv')
departments <- fread('E:/instacart/departments.csv')

all_tables = list('orders' = orders, 'products' = products, 'order_products'= order_products,'order_products_prior'= order_products_prior,'aisles' = aisles, 'departments' = departments)

```

```{r}
glimpse(orders)
glimpse(order_products_prior)
```


```{r}
two_tables = c(orders, order_products_prior)
class(all_tables)
class(two_tables)
#sapply(two_tables, glimpse) it is not working as expected
```

```{r}
#sapply(all_tables, kable(head())
for (a in all_tables) {
  b = kable(head(a,10))
  print(b)
}
#sapply(all_tables, kable(head()))
#lapply(all_tables, function(x) for (s in all_tables) {kable(head(s,10))})

```

```{r}
sapply(all_tables, dim)
sapply(all_tables, detach)
#sapply(two_tables, dim)# diff between list and c()?

```

```{r}
colnames(orders)
```

# We should do some recoding and convert character variables to factors
```{r}
for (x in all_tables){
  print(glimpse(x))
}

orders <- orders %>% mutate( eval_set = as.factor(eval_set), order_hour_of_day = as.factor(order_hour_of_day), order_dow = as.factor(order_dow))
products <- products %>% mutate( product_name = as.factor(product_name))
aisles <- aisles %>% mutate(aisle = as.factor(aisle))
departments <- departments %>% mutate(department = as.factor(department))

```

# When do people order (hour of the day)
```{r}

class(orders$order_hour_of_day)
orders %>%  ggplot(aes(x= order_hour_of_day)) + geom_histogram(stat = 'count', fill = 'blue' )

#aa = orders %>%  group_by(order_hour_of_day) %>% summarise(Freq_of_order_by_time = n()) %>% arrange(-Freq_of_order_by_time)
#plot(aa) why it is not coming in ascending
```

# Day of Week
```{r}
class(order_dow)
orders %>% ggplot(aes(x = order_dow)) + geom_histogram(stat = 'count' , fill = 'blue')
```

# When do they order again?
```{r}
fun_hist <-  function(df){
  list_col =  colnames(df)
  aa = list_col[sapply(list_col, function(x) is.factor(orders[,x]))]
  for (col in aa){
    plott = df %>% ggplot(aes_string(x = col)) + geom_histogram(stat = 'count' , fill = 'blue')
    print(plott)
    print(aa)
  }
}

fun_hist(orders)
```

```{r}

```


```{r}
kable(head(orders, 10))
```

# How many prior orders are there?
```{r}
#]orders %>% filter(eval_set=="prior") %>% count(order_number) %>% ggplot(aes(order_number,n)) + geom_line(color="red", size=1)+geom_point(size=2, color="red")
orders %>% filter(eval_set=="prior") %>% group_by(order_number) %>% summarise(n = NROW(order_number)) %>%  ggplot(aes(order_number,n)) + geom_line(color = 'blue') + geom_point(size = 1)

```


#How many items do people buy in one order?
```{r}
order_products %>% group_by(order_id) %>% summarise( n_items = n()) %>% ggplot(aes(x=n_items))+
  geom_histogram(stat = 'count', fill = 'blue') 
```

```{r}
#allaa = merge(orders,order_products_prior, by = 'order_id', all = F)
#View(all)
```

#Let's have a look which products are sold most often (top10)
```{r}
names(all_tables)
View(products)

freq_p = order_products %>% group_by(product_id) %>% summarise(freq_product = n())%>% arrange(-freq_product) %>% head(10)

freq_p

products$product_name[ products$product_id %in% freq_p$product_id ]

tmp <- products %>% filter(product_id %in% freq_p$product_id)

tmp
freq_p %>% left_join(select(products,product_id,product_name), by ="product_id")

```

# How often do people order the same items again?

```{r}
dsd <- order_products %>% 
  group_by(reordered) %>% 
  summarize(count = n()) %>% mutate(proportion = count/sum(count))
dsd
```

```{r}
dsd %>% ggplot(aes(x= reordered, y = count,fill = as.factor(reordered))) + geom_bar(stat = 'identity')
```

#Most often reordered


```{r}
tmp <-order_products %>% 
  group_by(product_id) %>% 
  summarize(proportion_reordered = mean(reordered), n=n()) %>% 
  filter(n>40) %>% 
  top_n(10,wt=proportion_reordered) %>% 
  arrange(desc(proportion_reordered)) %>% 
  left_join(products,by="product_id")

kable(tmp)
```


```{r}
tmp <-order_products %>% 
  group_by(product_id) %>% 
  summarize(proportion_reordered = mean(reordered), n=n(), prod = proportion_reordered*n) %>% top_n(10 , wt= prod) %>%  select(-prod) %>% left_join(products, by= "product_id")

tmp
```

#Which item do people put into the cart first?

```{r}
order_products %>% filter(add_to_cart_order == 1) %>% group_by( product_id) %>% summarise(countt = n()) %>% top_n(10 , wt = countt)
```

```{r}
aa = order_products %>% group_by( product_id, add_to_cart_order) %>% summarise(countt = n()) %>% arrange(add_to_cart_order, -countt) %>% group_by(product_id) %>% 
  mutate(total_addto =sum(countt)) %>% mutate( prop = countt / total_addto) %>% arrange(add_to_cart_order,-prop) %>% filter(countt > 10) %>% left_join(products, by = "product_id") 

aa = as.data.frame(aa)
plot_for_prop= aa %>% select(product_name, prop) %>% head(10)
plot_for_prop
plot_for_prop %>% ggplot(aes(x = reorder(product_name,-prop), y = prop)) + geom_bar(stat = 'identity', fill= "blue") + theme(axis.text.x=element_text(angle=90, hjust=1),axis.title.x = element_blank())+coord_cartesian(ylim=c(0.4,0.7))

```

#Association between time of last order and probability of reorder
```{r}
order_products %>% 
  left_join(orders,by="order_id") %>% 
  group_by(days_since_prior_order) %>%
  summarize(mean_reorder = mean(reordered)) %>%
  ggplot(aes(x=days_since_prior_order,y=mean_reorder))+
  geom_bar(stat="identity",fill="red")

```


#Association between number of orders and probability of reordering
```{r}
order_products %>% 
  group_by(product_id) %>% 
  summarize(proportion_reordered = mean(reordered), n=n()) %>%
  ggplot(aes(x=n,y=proportion_reordered))+
  geom_point()+
  geom_smooth(color="red")+
  coord_cartesian(xlim=c(0,2000))
```

#Organic vs Non-organic
```{r}

products <- products %>% 
    mutate(organic=ifelse(str_detect(str_to_lower(products$product_name),'organic'),"organic","not organic"), organic= as.factor(organic))
    
tmp <- order_products %>% 
  left_join(products, by="product_id") %>% 
  group_by(organic) %>% 
  summarize(count = n()) %>% 
  mutate(proportion = count/sum(count))
kable(tmp)
```




